Determining the alternatives for scalar implicature

### 2016-01-27 Paper/Paper.tex
reporting ME models, lot's a little tweaks

\begin{abstract}
Succesful communication regularly requires listeners to make pragmatic
inferences - enrichments beyond the literal meaning of a speaker's
utterance. For example, when interpreting a sentence such as ``Bob ate
some of the cookies,'' listeners routinely infer that Bob did not eat
all of them. A Gricean account of this phenomena assumes the presence of
alternatives (like ``all of the cookies'') with varying degrees of
informativity, but it remains an open question precisely what these
alternatives are. We use a computational model of pragmatic inference to
test out hypotheses about how well different sets of alternatives allow
us to predict scalar implicature performance across a range of different
scales. Our findings suggest that human comprehenders likely consider a
much broader set of alternatives beyond those entailed by the initial
description.

@@

How much of what we mean comes from the words that go unsaid? As
listeners, the ability to enrich meaning in context is an indispensible
component of succesful communication. For example, listeners commonly
enrich the meaning of the scalar item \emph{some} to \emph{some but not
all} in sentences like ``Alice ate some of the cookies'' (Grice, 1975;
Horn, 1984; Levinson, 2000). These inferences, called \emph{scalar
implicatures}, have been an important test case for understanding
pragmatic inferences more generally. A Gricean account of this phenomena
assumes listeners reason about the meaning the speaker intended by
incorporating knowledge about a) alternative scalar items a speaker
could have used (such as \emph{all}) and b) the relative informativity
of using such alternatives (Grice, 1975). According to this account, a
listener will infer that the speaker must have intended that Alice did
not eat \emph{all} the cookies because it would have been
\emph{underinformative} for the speaker to use \emph{some} when
\emph{all} could have been used.

But how much of the meaning of \emph{some} comes from the alternatives
that weren't chosen? Put differently, what are the alternatives that
should be considered in this computation more generally? Under classic
accounts of implicature, listeners consider only those words whose
meaning would entail the word that is actually sent (Horn, 1972), and
these alternatives enter into conventionalized or semi-conventionalized
scales (Levinson, 2000). For example, because \emph{all} entails
\emph{some}, and hence is a ``stronger'' meaning, \emph{all} should be
considered as an alternative to \emph{some} in implicatures. Similar
scales exist for non-quantifier scales, e.g. \emph{loved} entails
\emph{liked} (and hence ``I liked the movie'' implicates that I didn't
love it).

@@

\noindent In other words, the probability of a particular meaning given
an utterance is proportional to the speaker's proability of using that
particular utterance to express that meaning, weighted by a prior over
meanings. This prior represents the listener's \emph{a priori}
expectations about plausible meanings, independent of the utterance.
Because our experiments take place in a context in which listeners
should have very little expectation about which meanings speakers want
to convey, for simplicity we assume a uniform prior \(p(m) \propto 1\).

@@

\[p(u \mid m) \propto e^{-\alpha(-log(p_{L_0}(m \mid u)) - C(u))},\]

@@

To instantiate our signaling game with a tractable message set \(M\), in
our studies we adopt a food-review paradigm: we assume that speakers and
listeners are trying to communicate the number of stars in an onine
restaurant review (where \(m \in \{1, 2, 3, 4, 5\}\)). We then use
experiments to measure three components of the model. First, to measure
literal semantics \({p_{L_0} (m \mid u)}\) (we ask experiment
participants to judge whether a message is applicable to a particular
meaning (Experiment 1). Second, to generate a set of plausible
alternative messages in \(U\), we elicit alternatives directly
(Experiment 2). Lastly, to obtain human \(L_1\) pragmatic judgments, we
ask participants to interpret a speaker's utterances (Experiment 3).

@@

In each condition we recruited 30 participants from Amazon Mechanical
Turk (AMT). In the two-alternative condition, two participants were
excluded because they reported having a native language other than
English for a final sample of 28 participants. In the four-alternative
condition, six participants were excluded for either failing to pass two
training trials a non-English native language, leaving a total sample of
24 participants. In the five-alternative condition, there were no data
exclusions due to training failures or native language requirements,
leaving a total sample of 30 participants.

@@

Figure 2 plots literal listener \(p_{L0}(m|u)\) distributions for the
three conditions. Several trends are visible. First, in each scale, the
alternatives spanned the star scale, such that there were alternatives
that were highly compatible with both the lowest and highest numbers of
stars. Second, there was clear variability between scalar families. For
example, compatibility judgments for the top items in the
\emph{memorable / unforgettable} scale were more similar than those for
\emph{good / excellent} or \emph{liked / loved.} Finally, there was
substantial consistency in ratings for items that were repeated across
experiments, suggesting that this paradigm elicited stable judgments
from participants.

@@

\caption[Pragmatic listener judgements for our entailment scalar items]{Pragmatic listener judgements for our entailment scalar items. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Each line-type shows a different experiment, and colors indicate the stronger (ie excellent) or weaker (ie good) scalar items for each scale. Each panel shows one scalar pair.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

@@

\caption[Caption goes here]{Caption goes here}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

Figure \ref{fig:exp2_altsPlot} an example alternative set for the scalar
items \emph{liked} and \emph{loved} (combined). Alternative
distributions for the other scalar pairs (e.g.. \emph{good/excellent},
\emph{memorable/unforgettable}) were similarly long-tailed.

@@

We recruited 100 participants from AMT, 50 for each condition. Data for
7 participants in the two-alternative condition were excluded after
participants either failed to pass two training trials or reported a
non-English native language, leaving a total sample of 43 participants.
In the four-alternative condition, data from five participants was
thrown out after participants either failed to pass two training trials
or were not native English speakers, leaving a total sample of 45
participants.

@@

Participants were presented with a one-sentence prompt containing a
target scalar item such as ``Someone said they thought the food was
\_\_\_\_\_.'' Participants were then asked to generate a star rating
representing the rating they thought the reviewer likely gave. Each
participant was presented with all scalar items in a random order. The
experimental setup is shown in Figure \ref{fig:expt1_stimuli}, right.

@@

@@

\% latex table generated in R 3.2.3 by xtable 1.8-0 package \% Tue Jan
26 20:47:20 2016

>>>>

\begin{abstract}
Succesful communication regularly requires listeners to make pragmatic
inferences - enrichments beyond the literal meaning of a speaker's
utterance. For example, when interpreting a sentence such as ``Alice ate
some of the cookies,'' listeners routinely infer that Alice did not eat
all of them. A Gricean account of this phenomena assumes the presence of
alternatives (like ``all of the cookies'') with varying degrees of
informativity, but it remains an open question precisely what these
alternatives are. We use a computational model of pragmatic inference to
test hypotheses about how well different sets of alternatives allow us
to predict scalar implicature performance across a range of different
scales. Our findings suggest that human comprehenders likely consider a
much broader set of alternatives beyond those entailed by the initial
description.

@@

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is an indispensable component of successful
communication. For example, listeners commonly enrich the meaning of the
scalar item \emph{some} to \emph{some but not all} in sentences like
``Alice ate some of the cookies'' (Grice, 1975; Horn, 1984; Levinson,
2000). These inferences, called \emph{scalar implicatures}, have been an
important test case for understanding pragmatic inferences more
generally. A Gricean account of this phenomena assumes listeners reason
about the meaning the speaker intended by incorporating knowledge about
a) alternative scalar items a speaker could have used (such as
\emph{all}) and b) the relative informativity of using such alternatives
(Grice, 1975). According to this account, a listener will infer that the
speaker must have intended that Alice did not eat \emph{all} the cookies
because it would have been underinformative for the speaker to use
\emph{some} when \emph{all} could have been used.

But what are the alternatives that should be considered in this
computation more generally? Under classic accounts of implicature,
listeners consider only those words whose meaning would entail the word
that is actually sent (Horn, 1972), and these alternatives enter into
conventionalized or semi-conventionalized scales (Levinson, 2000). For
example, because \emph{all} entails \emph{some}, and hence is a
``stronger'' meaning, \emph{all} should be considered as an alternative
to \emph{some} in implicatures. Similar scales exist for non-quantifier
scales, e.g. \emph{loved} entails \emph{liked} (and hence ``I liked the
movie'' implicates that I didn't love it).

@@

\noindent In other words, the probability of a particular meaning given
an utterance is proportional to the speaker's probability of using that
particular utterance to express that meaning, weighted by a prior over
meanings. This prior represents the listener's \emph{a priori}
expectations about plausible meanings, independent of the utterance.
Because our experiments take place in a context in which listeners
should have very little expectation about which meanings speakers want
to convey, for simplicity we assume a uniform prior \(p(m) \propto 1\).

@@

\[p(u \mid m) \propto e^{-\alpha(-log(p_{L_0}(m \mid u)) - C(u))}\]

@@

To instantiate our signaling game with a tractable message set \(M\), in
our studies we adopt a food-review paradigm: we assume that speakers and
listeners are trying to communicate the number of stars in an online
restaurant review (where \(m \in \{1, 2, 3, 4, 5\}\)). We then use
experiments to measure three components of the model. First, to measure
literal semantics \({p_{L_0} (m \mid u)}\) (we ask experiment
participants to judge whether a message is applicable to a particular
meaning (Experiment 1). Second, to generate a set of plausible
alternative messages in \(U\), we elicit alternatives directly
(Experiment 2). Lastly, to obtain human \(L_1\) pragmatic judgments, we
ask participants to interpret a speaker's utterances (Experiment 3).

@@

In each condition we recruited 30 participants from Amazon Mechanical
Turk (AMT). In the two-alternative condition, 16 participants were
excluded
\footnote{The majority of respondent data excluded from Experiment 1a was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in Experiments 1b,1c}
for either failing to pass two training trails or were not native
English speakers, leaving a total sample of 14 participants. In the
four-alternative condition, 7 participants were excluded for either
failing to pass two training trials or were not native English speakers,
leaving a total sample of 23 participants. In the five-alternative
condition, 3 participants were excluded for either failing to pass two
training trials or were not native English speakers, leaving a total
sample of 27.

@@

Figure \ref{fig:exp2Plots} plots literal listener \(p_{L0}(m|u)\) scalar
item distributions for the three conditions. Each row plots a unique
scalar family with items ordered horizontally by valence. Several trends
are visible. First, in each scale, the alternatives spanned the star
scale, such that there were alternatives that were highly compatible
with both the lowest and highest numbers of stars. Second, there was
clear variability between scalar families. For example, compatibility
judgments for the top items in the \emph{memorable / unforgettable}
scale were more similar than those for \emph{good / excellent} or
\emph{liked / loved.} Finally, there was substantial consistency in
ratings for items that were repeated across experiments, suggesting that
this paradigm elicited stable judgments from participants.

To test whether there were between experiment differences in
acceptability judgments, we ran a mixed effects model, regressing
acceptability judgements on scale, number of stars and experiment type,
with by subject and by word random effects. Results indicate no
significant differences between Experiment 1a and 1c \(b =\)-0.05,
\(Z =\) -0.53, \(p =\) 0.59 or between 1b and 1c \(b =\) -0.04, \(Z =\)
-0.52 \(p =\) 0.6 and the addition of our experiment predictor did not
significantly improve model fit when compared to a model without the
experiment type variable using ANOVA \(\chi^2(\) 2 \() =\) 0.43, \(p =\)
0.81.

@@

\caption[Pragmatic listener judgements for our entailment scalar items]{Pragmatic listener judgements for our entailment scalar items. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Each line-type shows a different experiment, and colors indicate the stronger (ie excellent) or weaker (ie good) scalar items for each scale. Each panel shows one scalar pair. We are only showing scalar items used for model comparison (items used in Experiments 1a / 3a) for simplicity,}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

@@

\caption[Combined counts for participant generated alternatives for 'liked  loved' scale]{Combined counts for participant generated alternatives for 'liked  loved' scale. Participants were shown a target scalar (either liked or loved) and asked to give three alternatives.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

Figure \ref{fig:exp2_altsPlot_likedLoved} an example alternative set for
the scalar items \emph{liked} and \emph{loved} (combined). Alternative
distributions for the other scalar pairs (e.g.. \emph{good/excellent},
\emph{memorable/unforgettable}) were similarly long-tailed.

@@

We recruited 100 participants from AMT, 50 for each condition. Data for
9 participants in the two-alternative condition were excluded after
participants either failed to pass two training trials or reported a
non-English native language, leaving a total sample of 41 participants.
In the four-alternative condition, data from 7 participants was thrown
out after participants either failed to pass two training trials or were
not native English speakers, leaving a total sample of 43 participants.

@@

Participants were presented with a one-sentence prompt containing a
target scalar item such as ``Someone said they thought the food was
\_\_\_\_\_.'' Participants were then asked to generate a star rating
representing the rating they thought the reviewer likely gave. Each
participant was presented with all scalar items in a random order. The
experimental setup is shown in Figure \ref{fig:stimuli_exp1}, right.

@@

To test whether there were between experiment differences in pragmatic
judgments, we ran a mixed effects model, regressing pragmatic judgements
on scale and experiment with subject and word level random effects.
Results indicate there were no significant differences between
Experiment 3a and 3b for \(b =\) 0.05, \(t(\) 150.4 \() =\) 1.04,
\(p =\) 0.3 and the addition of our experiment predictor did not
significantly improve model fit when compared to a model without the
experiment variable using ANOVA \(\chi^2(\) 1 \() =\) 1.13, \(p =\)
0.29.

@@

\% latex table generated in R 3.2.3 by xtable 1.8-0 package \% Wed Jan
27 09:40:50 2016
